---
output: github_document
author: Joachim Gassen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(rdfanalysis)
library(knitr)
```

# Researcher Degrees of Freedom Analysis

## A package to explore and document your degrees of freedom

This experimental in-development package provides a set of functions 
to develop data anylsis code that systematically documents
researcher degrees of freedom when conducting analyses on observational data. 
The resulting code base is self-documenting, 
supports unit testing and power simulations based on simulated data. The 
documented researcher degrees of freedom can be exhausted to generate a 
distribution of outcome estimates.

### First: Define your research design by a series of functions
    
```{r steps}
sim_data <- function(input = NULL, choice = NULL) {
  step_description <- doc(
    "## Simulates Data",
    "### Content",
    "",
    "This step simulates some data.",
    "The used equation is $y = x + z + \\epsilon$",
    "with $z$ being N(0,1) distributed and $x$ being potentially endogenous to $z$ (see choices below)"
  )
  choice_description <- doc(
    "### Choice",
    "",
    "A list containing a character value `make_x_endogenous` and a numerical value `noise_y`.", 
    "`make_x_endogenous` may take one of the following values:",
    "",
    "- `yes`: x will depend on z",
    "- `no`: x will not depend on z",
    "",
    "`noise_y` sets the standard deviation of the error term affecting y",
    "and may take any value within [1, 2]"
  )
  choice_type <- list(
    list(name = "make_x_endogenous", 
         type = "character", 
         valid_values = c("yes", "no")),
    list(name = "noise_y", 
         type = "double", 
         valid_min = 1, valid_max = 2)
  )
  if (is.null(choice)) return(list(
    step_description = step_description,
    choice_description = choice_description,
    choice_type = choice_type
  )) else check_choice(choice, choice_type)

  z <- rnorm(1000)
  x <- rnorm(1000) 
  if (choice[[1]] == "yes") x <- x + z
  y <- x + z + rnorm(1000, sd = choice[[2]]) 
  return(list(
    data = data.frame(x = x, y = y, z = z),
    protocol = choice
  ))
}


est_model <- function(input = NULL, choice = NULL) {
  step_description <- doc(
    "## Estimate model",
    "### Content",
    "",
    "This step estimates on OLS model based on the simulated data of the last step"
  )
  choice_description <- doc(
    "### Choice",
    "",
    "A character value `control_for_z` that may take one of the following values:",
    "",
    "- `yes`: control for z",
    "- `no`: do not control for z"
  )
  choice_type <- list(
    list(name = "control_for_z", 
         type = "character", 
         valid_values = c("yes", "no"))
  )
  if (is.null(choice)) return(list(
    step_description = step_description,
    choice_description = choice_description,
    choice_type = choice_type
  )) else check_choice(choice, choice_type)

  if(choice[[1]] == "yes") 
    mod <- lm(y ~ x + z, data = input$data)
  else     mod <- lm(y ~ x, data = input$data)
  return(list(
    data = list(
      est = summary(mod)$coefficient[2,1],
      lb = confint(mod)[2,1],
      ub = confint(mod)[2,2]
    ),
    protocol = c(input$protocol, choice)
  ))  
} 

design <- c("sim_data", "est_model")
```

## Second: Display your design and verify it by testing it

```{r test_design}
prepare_design_flow_chart(design, landscape = TRUE)

test_design(design, reporter = "minimal")
```

# Third: Run a single protocol

```{r single_protocol}
sim_data(NULL, list("no", 2)) %>%
  est_model("no")
```

# Fourth: Exhaust your research degrees of freedom

```{r exhaust_rdf, results = "hide"}
df <- exhaust_design(design, NULL) 
```

```{r display_rdf}
plot_rdf_estimates(df, est = "est", lb = "lb", ub = "ub")
kable(df)
```
